Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-02-04 paper/main.tex
minor edit

Combinatorial Efficient Global Optimization (CEGO)~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO
method~\citep{JonSchWel98go} to unconstrained black-box combinatorial
optimization problems. In EGO, Gaussian process models are used as a surrogate
of the landscape of the expensive original problem. An optimization method
searches for solutions in the surrogate model by optimizing the expected
improvement criterion, which balances the expected mean and variance of the
chosen solution. Once a solution is chosen, it is evaluated on the actual
fitness function and the result is used to update the surrogate-model,
hopefully increasing its predictive power.

CEGO replaces the Euclidean distance measure, used by the surrogate model in
EGO, with a distance measure appropriate to combinatorial
landscapes~\citep{ZaeStoBar2014:ppsn}, such as Kendall's-$\tau$ distance for
permutations (Eq.~\ref{eq:kendall_decomp}). In CEGO, the surrogate model is
explored by a GA with crossover and mutation operators appropriate for the
particular combinatorial problem. The original paper notes that coupling the GA
with local search does not improve the results significantly since the model is
anyway an inexact estimation of the original fitness
function~\citep[p.~875]{ZaeStoFriFisNauBar2014}. A thorough comparison of CEGO
with other optimizers for expensive black-box combinatorial optimization ranked
it as the best-performing~\citep{ZaeStoFriFisNauBar2014}.

>>>>

\myparagraph{CEGO}
%
Combinatorial Efficient Global Optimization (CEGO)~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO
method~\citep{JonSchWel98go} to unconstrained black-box combinatorial
optimization problems. In EGO, Gaussian process models are used as a surrogate
of the landscape of the expensive original problem. An optimization method
searches for solutions in the surrogate model by optimizing the expected
improvement criterion, which balances the expected mean and variance of the
chosen solution. Once a solution is chosen, it is evaluated on the actual
fitness function and the result is used to update the surrogate-model,
hopefully increasing its predictive power.

CEGO replaces the Euclidean distance measure, used by the surrogate model in
EGO, with a distance measure appropriate to combinatorial
landscapes~\citep{ZaeStoBar2014:ppsn}.
% , such as Kendall's-$\tau$ distance for permutations
% (Eq.~\ref{eq:kendall_decomp}).
In CEGO, the surrogate model is
explored by a GA with crossover and mutation operators appropriate for the
particular combinatorial problem. The original paper notes that coupling the GA
with local search does not improve the results significantly since the model is
anyway an inexact estimation of the original fitness
function~\citep[p.~875]{ZaeStoFriFisNauBar2014}. A thorough comparison of CEGO
with other optimizers for expensive black-box combinatorial optimization ranked
it as the best-performing~\citep{ZaeStoFriFisNauBar2014}.
